---
title: "NB_PGPR+PVY"
author: "Bibechana Paudel"
date: "2025-04-14"
output:
  md_document:
    variant: gfm
  html_document:
    toc: TRUE
    toc_float: TRUE
---


# First trial

###Libraries used

```{r}
library(tidyverse)
library(ggplot2)
library(agricolae)
library(tidyr)
library(ggpubr)
library(multcomp)
library(emmeans)
library(multcompView)
```

###Load and verify data of Rep 1
```{r}
rep1 <- read.csv("Nb_PGPR+PVY_1st_Reproducibility.csv", header=T)
head(rep1)  # View the first few rows
str(rep1)   # Check the structure of the data
summary(rep1) # Quick statistical summary
```

###Make the factor variable
```{r}
rep1$Treatment<-as.factor(rep1$Treatment)
rep1$Dpi<-as.factor(rep1$Dpi)
```

##Calculate the ViralLoad using the formula derived by .... and log transform for normality. 
```{r}
rep1<-rep1 %>% 
mutate (ViralLoad=((10^((Cq-49.153)/-3.93)))) %>% 
mutate(logViralLoad=log10(ViralLoad))
```

## Statistical analysis

```{r}
model<-lm(logViralLoad~Treatment*Dpi,data=rep1)
car::Anova(model)

lsmeans <- emmeans(model, ~Treatment|Dpi) # estimate lsmeans of variety within siteXyear
Results_lsmeansEC <- multcomp::cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE,  Letters = letters) # contrast with Tukey ajustment
Results_lsmeansEC
```

##Data summarization

```{r}
# Extracting the letters for the bars
sig.diff.letters <- data.frame(Results_lsmeansEC$emmeans$Treatment, 
                               Results_lsmeansEC$emmeans$Dpi,
                               str_trim(Results_lsmeansEC$emmeans$.group))
colnames(sig.diff.letters) <- c("Treatment", 
                                "Dpi",
                                "Letters")
```

##Extract significance letters for plotting

```{r}
PVY1 <- rep1 %>%
  group_by(Treatment, Dpi) %>%
  dplyr::summarize(
    avg.virus= mean(logViralLoad, na.rm=TRUE),
    n=n(),
    se = sd(logViralLoad)/sqrt(n)) %>%
  left_join(sig.diff.letters)
```

##Classify the sample collection days based on leaf

```{r}
PVY1$Leaf_sample <- ifelse(
  grepl("^IL", PVY1$Dpi), "Inoculated",  # Matches Dpi values starting with IL
  ifelse(grepl("^SL", PVY1$Dpi), "Systemic", NA)
)
```

## Visualization: Inoculated sample

```{r}
inoculated<-filter(PVY1, Dpi=="IL-1" | Dpi=="IL-4" | Dpi=="IL-7" | Dpi=="IL-10")
inoculated$Treatment<-factor(inoculated$Treatment, levels=c("Control","P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
```


```{r PVY on IL_1st trial_NB }
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00", "#009E73")
PVY_IL_1<-ggplot(inoculated, aes(x =factor(Dpi,levels=c("IL-1", "IL-4", "IL-7", "IL-10")), y =avg.virus , fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  geom_text(data = inoculated, aes(label = Letters, y = avg.virus),position = position_dodge(width = 0.7), vjust = -0.5)+
  theme_classic() +
  labs(title= "" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 8),
               axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )

 PVY_IL_1 
```

## Visualization: Systemic sample

```{r}
systemic<-filter(PVY1, Dpi=="SL-7" | Dpi=="SL-10")
systemic$Treatment<-factor(systemic$Treatment, levels=c("Control","P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
```


```{r, PVY.systemic.1sttrial}
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00", "#009E73")
PVY_SL_1<-ggplot(systemic, aes(x =factor(Dpi,levels=c("SL-7","SL-10")), y =avg.virus , fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  geom_text(data = systemic, aes(label = Letters, y = avg.virus),position = position_dodge(width = 0.7), vjust = -0.5)+
  theme_classic() +
  labs(title= "" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 8),
                axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )

 PVY_SL_1 
```

## Combined figure

```{r}
PVY.1st<-ggarrange(PVY_IL_1, PVY_SL_1,
                   nrow=1,
                   ncol=2,
                   common.legend=T,
                   widths=c(1.75,1))
PVY.1st
```



# Second trial

##Load and verify the data

```{r}
rep2 <- read.csv("Nb+PGPR+PVY_2nd_Reproducibility.csv", na.strings="na")
str(rep2)   # Check the structure of the data
summary(rep2) # Quick statistical summary
```

###Make the factor variable
```{r}
rep2$Treatment<-as.factor(rep2$Treatment)
rep2$Dpi<-as.factor(rep2$Dpi)
rep2$ViralLoad<-((10^((rep2$Cq-49.153)/-3.93)))
```

```{r}
# Reorder the Treatment  and Dpi (IL only) factor
rep2$Treatment <- factor(rep2$Treatment, levels = c("Control", "P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
rep2_IL <- rep2[rep2$Dpi %in% c("IL-1", "IL-4", "IL-7", "IL-10"), ]
```


```{r, IL-Rep2}
####Plot Bar graph
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00", "#009E73")
PVY2_IL<-ggplot(rep2_IL, aes(x =factor(Dpi,levels=c("IL-1", "IL-4", "IL-7", "IL-10")), y = log10(ViralLoad), fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title= "IL" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.3)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 7),
        plot.title= element_text(hjust = 0.5), # Center the plot title
        axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )
PVY2_IL
```



```{r}
# Reorder the Treatment and Dpi (SL only) factor
rep2$Treatment <- factor(rep1$Treatment, levels = c("Control", "P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
rep2_SL <- rep2[rep2$Dpi %in% c("SL-7", "SL-10"), ]
```


```{r, SL-Rep2}
####Bar graph
PVY2_SL<-ggplot(rep2_SL, aes(x =factor(Dpi,levels=c("SL-7", "SL-10")), y = log10(ViralLoad), fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title= "SL" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.3)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 7),
        plot.title= element_text(hjust = 0.5), # Center the plot title
        axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
        
  )
PVY2_SL
```

###Loop for subsetting the Dpi and linear modeling

```{r}
# Create vector of cytokines to analyze
R2 <- c("IL-1", "IL-4", "IL-7", "IL-10","SL-7","SL-10")

# Loop through cytokines and print model summaries
for (dpi in R2) {
  cat("\nModel Summary for", dpi, ":\n")
  
  rep2 %>%
    filter(Dpi == dpi) %>%
    lm(log10(ViralLoad) ~ Treatment + Replication, data = .) %>%
    summary() %>%
    print()
}
```

# Third trial

```{r}
## Load data Rep 3
rep3 <- read.csv("Nb_PGPR+PVY_3rd_Reproducibility.csv", na.strings="na")
####Verify the data
head(rep3)  # View the first few rows
str(rep3)   # Check the structure of the data
summary(rep3) # Quick statistical summary
```

###Make the factor variable
```{r}
rep3$Treatment<-as.factor(rep3$Treatment)
rep3$Dpi<-as.factor(rep3$Dpi)
```


```{r}
rep3$ViralLoad<-((10^((rep3$Cq-49.153)/-3.93)))
```


```{r}
# Reorder the Treatment and Dpi (IL only) factor
rep3$Treatment <- factor(rep3$Treatment, levels = c("Control", "P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
rep3_IL <- rep3[rep3$Dpi %in% c("IL-1", "IL-4", "IL-7", "IL-10"), ]
```


```{r, IL-Rep3}
####Bar graph
cbbPalette <- c( "#56B4E9","#F0E442","#CC79A7", "#E69F00", "#009E73")
PVY3_IL<-ggplot(rep3_IL, aes(x =factor(Dpi,levels=c("IL-1", "IL-4", "IL-7", "IL-10")), y = log10(ViralLoad), fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title= "IL" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.3)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 7),
        plot.title= element_text(hjust = 0.5), # Center the plot title
        axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
  )
PVY3_IL
```


```{r}
# Reorder the Treatment and Dpi (SL only) factor
rep3$Treatment <- factor(rep3$Treatment, levels = c("Control", "P. fluorescens", "S. marcescens", "B. amyloliquifaciens", "B. subtilis"))
rep3_SL <- rep3[rep3$Dpi %in% c("SL-7", "SL-10"), ]
```


```{r, SL-Rep3}
####Bar graph

PVY3_SL<-ggplot(rep3_SL, aes(x =factor(Dpi,levels=c("SL-7", "SL-10")), y = log10(ViralLoad), fill = Treatment)) +
  stat_summary(fun=mean,geom="bar", position = position_dodge(width = 0.7), width = 0.6) +
  stat_summary(fun.data = mean_se, geom="errorbar",width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title= "SL" ,
       x = "",
       y = "Log Copy Number") +
  scale_fill_manual(values=cbbPalette,
                    labels=c("Control",expression(italic("P. fluorescens")), 
                             expression(italic("S. marcescens")),expression(italic("B. amyloliquifaciens")),expression(italic("B. subtilis"))))+
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.3)), breaks = seq(1, 9, by = 1)  # This removes padding at the bottom of the plot
  )+
  theme(legend.position = "right",
        legend.key.size = unit(0.5, "cm"),  # Adjusts the size of the legend keys
        legend.title = element_blank(),  # Adjusts the legend title font size
        legend.text = element_text(size = 7),
        plot.title= element_text(hjust = 0.5), # Center the plot title
        axis.text.x = element_text(color = "black", face="bold"),   # X-axis text color
        axis.text.y = element_text(color = "black", face="bold"),   # Y-axis text color
        axis.title.x = element_text(color = "black"),  # X-axis label color
        axis.title.y = element_text(color = "black", face="bold", size=12),  # Y-axis label color
        axis.line = element_line(color = "black")      # Axis lines color
        
  )
PVY3_SL

```

###Loop for subsetting the Dpi and linear modeling

```{r}
# Create vector of cytokines to analyze
R3 <- c("IL-1", "IL-4", "IL-7", "IL-10","SL-7","SL-10")

# Loop through cytokines and print model summaries
for (dpi in R3) {
  cat("\nModel Summary for", dpi, ":\n")
  
  rep3 %>%
    filter(Dpi == dpi) %>%
    lm(log10(ViralLoad) ~ Treatment + Replication, data = .) %>%
    summary() %>%
    print()
}
```



